---
title: "Seurat preprocessing"
author:
    - "Chia-Chen Chu"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
  highlight: tango
  number_sections: no
  theme: default
  toc: yes
  toc_depth: 3
  toc_float:
    collapsed: no
    smooth_scroll: yes
    
editor_options: 
  markdown: 
    wrap: 72
---

NOTE: Run with docker image chiaenu/seurat-rmd:5.0.0

# Input data

Gene expression matrix
`GSM5453540_filtered_feature_bc_matrix_Lch.tar.gz` were downloaded from
NCBI GEO as `data/lch/GSM5453540_filtered_feature_bc_matrix_Lch.tar.gz`

Extract data using `tar -xvzf` gives following files:

    filtered_feature_bc_matrix/
    filtered_feature_bc_matrix/features.tsv.gz
    filtered_feature_bc_matrix/matrix.mtx.gz
    filtered_feature_bc_matrix/barcodes.tsv.gz

# Setups

Import libraries

```{r}
library(Seurat)
library(magrittr)
```

Set args

```{r}
sample_name <- "lch"
input_file <- sprintf("data/lch/filtered_feature_bc_matrix")
output_dir <- sprintf("outputs/Seurat/%s", sample_name)

# create output directory
dir.create(output_dir, recursive = TRUE, showWarnings = FALSE) 

# setup output file paths
path_rdata <- file.path(output_dir, "seurat_obj.rds") 
path_pca <- file.path(output_dir, "pca.csv") 
path_clusters <- file.path(output_dir, "clusters.csv")
```

# Analysis

Read data

```{r}
seurat_obj <- Read10X(data.dir = input_file) %>% CreateSeuratObject()
seurat_obj
```

Normalize data with sctransform

```{r}
seurat_obj <- SCTransform(seurat_obj, variable.features.n = 2000)
VariableFeaturePlot(seurat_obj)
```

Run PCA

```{r}
seurat_obj <- RunPCA(seurat_obj, npcs = 100, verbose = TRUE)
ElbowPlot(seurat_obj, ndims = 100, reduction = "pca")
```

Set the number of PC for downstream analyses

```{r}
npcs <- 30
```

Clustering

(adjust the clustering resolution so that the number of clusters matches
that in the original study)

```{r}
seurat_obj <- FindNeighbors(seurat_obj, reduction = "pca", dims = 1:npcs)
seurat_obj <- FindClusters(seurat_obj, resolution = .4)
table(seurat_obj$seurat_clusters)
```

# Export data

```{r}
saveRDS(seurat_obj, file = path_rdata)
write.csv(
  Embeddings(seurat_obj, reduction = "pca")[, 1:npcs], 
  file = path_pca, 
  quote = FALSE
)
write.csv(
  Idents(seurat_obj) %>% as.data.frame,
  file = path_clusters,
  quote = FALSE
)
```

Print session info 

```{r}
sessionInfo()
```